Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-01-19 papers/clusters.tex
biblio tweaks

% ARC to DWH: Are the colors of broken glass indicative of anything?

@@

The abundances of known clusters that we identify by k-means are in
excellent agreement with the literature. For the group shown in Figure \ref{fig:M13},
which we later attribute to M13, we find a mean of [Fe/H]$ = -1.55$ (DWH CHECK).
The mean and spread agree well with other works dedicated to studying this cluster
\citep{Kraft_1992,Cohen_Melendez_2005,Johnson_Pilachowski_2012}, and compiled catalogs of
globular cluster properties \citep[][accessed 2016]{Harris_1996}. The detailed abundances are also
consistent: we find a correlation in [C/Fe] and [N/Fe] abundances \citep{Smith_2005}, and their
projection in Figure \ref{fig:M13} show a spread that reflects deep mixing
along the red giant branch \citep{Briley_2004}. We find that the light element
abundances (most notably Mg--Al) for this cluster are anti-correlated, as
expected from other studies \citep[e.g.,][]{gratton}. However we find a
smaller spread in [O/Fe] than reported by others who have looked exclusively
at M13 \citep{Johnson_Pilachowski_2012}.  As we discuss above, this is expected:
k-means is by no means optimal for identifying arbitrary-shaped structures
in high-dimensional space, and the first stars that would be `discarded' from
this group would be those with the most extreme abundances: low [O/Fe] and
high [Na/Fe].

There are scant abundance measurements reported for Palomar 5, for which
most stars in the cluster in Figure \ref{fig:Pal5} belong. We find a mean
metallicity of [Fe/H]$ = -1.3$ (DWH CHECK), which lies between the only
photometric \citep{Geisler_1997} and spectroscopic \citep{Smith_1985,Smith_2002} abundance
measurements for this cluster.  The anti-correlation in C--N and Na--O abundance
have been seen elsewhere \citep[][respectively]{Smith_1985,Smith_2002}, although
to our knowledge the Mg--Al anti-correlation has not previously been shown for this cluster.
Indeed, this work contains the largest number of abundances for Palomar 5,
permitting a thorough search of tidally disrupted members \citep[e.g.,][]{Kuzma_2015} 
that have retained the unique cluster abundance signature.

@@

 \bibitem[Briley \etal(2004)]{Briley_2004} Briley, M.~M., Cohen, 
J.~G., \& Stetson, P.~B.\ 2004, \aj, 127, 1579 
\bibitem[Casey \etal(2016)]{casey16}

@@

\bibitem[Cohen \& Mel{\'e}ndez(2005)]{Cohen_Melendez_2005} Cohen, J.~G., \& Mel{\'e}ndez, J.\ 2005, \aj, 129, 303 

@@

\bibitem[Geisler \etal(1997)]{Geisler_1997} Geisler, D., Claria, 
J.~J., \& Minniti, D.\ 1997, \pasp, 109, 799 

@@

\bibitem[Harris(1996)]{Harris_1996} Harris, W.~E.\ 1996, \aj, 112, 
1487 

@@

\bibitem[Johnson \& Pilachowski(2012)]{Johnson_Pilachowski_2012} Johnson, C.~I., \& Pilachowski, C.~A.\ 2012, \apjl, 754, L38 
\bibitem[Kraft \etal(1992)]{Kraft_1992} Kraft, R.~P., Sneden, C., Langer, G.~E., \& Prosser, C.~F.\ 1992, \aj, 104, 645 
\bibitem[Kuzma \etal(2015)]{Kuzma_2015} Kuzma, P.~B., Da Costa, 
G.~S., Keller, S.~C., \& Maunder, E.\ 2015, \mnras, 446, 3297 

@@

\bibitem[Smith(1985)]{Smith_1985} Smith, G.~H.\ 1985, \apj, 298, 
249 
\bibitem[Smith \etal(2002)]{Smith_2002} Smith, G.~H., Sneden, C., 
\& Kraft, R.~P.\ 2002, \aj, 123, 1502 
\bibitem[Smith \etal(2005)]{Smith_2005} Smith, G.~H., Briley, 
M.~M., \& Harbeck, D.\ 2005, \aj, 129, 1589

>>>>

@@

The abundances of known clusters that we identify by k-means are in
excellent agreement with the literature. For the group shown in Figure \ref{fig:M13},
which we later attribute to M13, we find a mean of [Fe/H]$ = -1.55$ (DWH CHECK).
The mean and spread agree well with other work on this cluster
\citep{Kraft_1992,Cohen_Melendez_2005,Johnson_Pilachowski_2012}, and compiled catalogs of
globular cluster properties \citep[][accessed 2016]{Harris_1996}. The detailed abundances are also
consistent: we find a correlation in [C/Fe] and [N/Fe] abundances \citep{Smith_2005}, and their
projection in Figure \ref{fig:M13} show a spread that reflects deep mixing
along the red giant branch \citep{Briley_2004}. We find that the light element
abundances (most notably Mg--Al) for this cluster are anti-correlated, as
expected from other studies \citep[for example,][]{gratton}. However we find a
smaller spread in [O/Fe] than reported by others who have looked exclusively
at M13 \citep{Johnson_Pilachowski_2012}.  As we discuss above, this is expected:
k-means is by no means optimal for identifying arbitrary-shaped structures
in high-dimensional space, and the first stars that would be `discarded' from
this group would be those with the most extreme abundances: low [O/Fe] and
high [Na/Fe].

There are scant abundance measurements reported for Palomar 5, for which
most stars in the cluster in Figure \ref{fig:Pal5} belong. We find a mean
metallicity of [Fe/H]$ = -1.3$ (DWH CHECK), which lies between the only
photometric \citep{Geisler_1997} and spectroscopic \citep{Smith_1985,Smith_2002} abundance
measurements for this cluster.  The anti-correlation in C--N and Na--O abundance
have been seen elsewhere \citep[][respectively]{Smith_1985,Smith_2002}, although
to our knowledge the Mg--Al anti-correlation has not previously been shown for this cluster.
Indeed, this work contains the largest number of abundances for Palomar 5,
permitting a thorough search of tidally disrupted members \citep[for example,][]{Kuzma_2015} 
that have retained the unique cluster abundance signature.

@@

 \bibitem[Briley \etal(2004)]{Briley_2004}
Briley, M.~M., Cohen, J.~G., Stetson, P.~B., 2004, \aj, 127, 1579 
\bibitem[Casey \etal(2016)]{ycasey16}

@@

\bibitem[Cohen \& Mel{\'e}ndez(2005)]{Cohen_Melendez_2005}
Cohen, J.~G.,  Mel{\'e}ndez, J., 2005, \aj, 129, 303 

@@

\bibitem[Geisler \etal(1997)]{Geisler_1997}
Geisler, D., Claria, J.~J., Minniti, D., 1997, \pasp, 109, 799 

@@

\bibitem[Harris(1996)]{Harris_1996}
Harris,~W.~E., 1996, \aj, 112, 1487 

@@

\bibitem[Johnson \& Pilachowski(2012)]{Johnson_Pilachowski_2012}
Johnson, C.~I.,  Pilachowski, C.~A., 2012, \apjl, 754, L38 
\bibitem[Kraft \etal(1992)]{Kraft_1992}
Kraft, R.~P., Sneden, C., Langer, G.~E.,  Prosser, C.~F., 1992, \aj, 104, 645 
\bibitem[Kuzma \etal(2015)]{Kuzma_2015}
Kuzma, P.~B., Da Costa, G.~S., Keller, S.~C., Maunder, E., 2015, \mnras, 446, 3297 

@@

\bibitem[Smith(1985)]{Smith_1985}
Smith, G.~H., 1985, \apj, 298, 249 
\bibitem[Smith \etal(2002)]{Smith_2002}
Smith, G.~H., Sneden, C., Kraft, R.~P., 2002, \aj, 123, 1502 
\bibitem[Smith \etal(2005)]{Smith_2005}
Smith, G.~H., Briley,~M.~M.,  Harbeck, D., 2005, \aj, 129, 1589
